### 2021-11-22 input/proposal.tex
meow

\begin{definition}[Autonomous systems]
  An autonomous system is a network comprising nodes and communication
  links that we represent as a \underline{g}raph of
  \underline{v}ertices and \underline{e}dges: $G = \langle V, E
  \rangle$ with $E \in V \times V$. A \underline{p}ath $\pi_{ij}$ from
  \Process~$i$ to \Process~$j$ is a sequence of vertices $[i, k_1,
    k_2, \ldots k_n, j]$ with $\forall m: 0\leq m \leq n, \langle
  \pi_{ij}[m], \pi_{ij}[m+1] \rangle \in E$.
\end{definition}

@@

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process~$x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process to
  \process: $b_x(m) \wedge r_y(m) \implies \exists \pi_{xy}: \forall z
  \in \pi_{xy}, \phi(\mu_z, \sigma_z)$.
\end{definition}

This definition encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}.  It
also highlights that epidemic propagation and scoped broadcast have
well-aligned preoccupations. More precisely, it underlines the
transitive relevance of messages, thus a \process can stop forwarding
messages as soon as its predicate becomes unverified.
%For instance, a \process
%from Paris could scoped broadcast messages to all \processes in
%Paris. This requires \processes to store and maintain their city
%location in their local state. \Processes stop delivering and
%forwarding their received messages when they come from a different
%city.  In another instance, a \process from Paris could scoped
%broadcast messages to all \processes in Paris plus neighboring
%cities. This requires \processes to overload forwarded messages the
%first time they reach another city. The predicate checks if messages
%already reached two distinct cities before delivery. Similarly to
%uniform reliable broadcast, scoped broadcast implementations expose
%different trade-offs on space, time, and communication.
%
We use \NAMEB to efficiently modify the state of each \process
depending on the partitions that exist in the system.

@@

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_{s\in
    S}$ be the \underline{p}artition including at least \Process~$s$,
  each \process belongs to at most one partition $\forall p,q \in V,
  \forall s,s' \in S: (p \in P_{s} \wedge q \in P_{s'}) \implies (p \neq
  q \vee s = s')$, and there exists at least one path $\pi_{ps}$ of
  \processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy}$ be the positive \underline{w}eight from $x$ to $y$,
  $\Pi_{xz}$ be the shortest \underline{p}ath from $x$ to $z$ the
  weight of which $|\Pi_{xz}|$ is lower than any other path weight,
  with $|\Pi_{xx}|$ being the greatest lower bound of $x$, the only
  consistent partitioning $\mathcal{P}$ is a set of partitions
  $P_{s\in S}$ such that each \process belongs to a logical partition
  comprising its closest source: $\forall p \in P_{s}: \nexists
  P_{s'}$ such that $|\Pi_{s'p}| < |\Pi_{sp}|$.
\end{definition}

@@

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $p$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors: $f_p(m) \implies \forall
  \langle p, q\rangle \in E: \TODO{s_{pq}(m)}$.
\end{definition}

\begin{theorem}[\label{theo:efb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually$CP]
  %
  Assuming reliable communication links ($s_{pq}(m) \iff \eventually
  r_{qp}(m)$), \processes eventually reach consistent partitioning if
  each \process delivers and forwards its best message among
  \underline{r}eceived messages $R_p$, \ie the message $\alpha_s^d$
  with the smallest distance $d$ to the source $s$ ($(\alpha_s^d = \min R_p)
  \implies (d_p(\alpha_s^d) \wedge f_p(\TODO{\alpha_s^d}))$).
\end{theorem}

@@

\begin{proof}
  When there are no sources, no \process receives, nor delivers hence
  forwards any message. \Processes do not belong to any partition.

  \noindent Whenever a \process $s$ becomes a source, it delivers its
  own message $d_p(\alpha_{s})$. Whatever its set of received
  messages, its own partition $P_s$ becomes and remains its closest
  partition $\alpha_{s} = \min D_p$ since $\forall p \neq s: |\Pi_{s
    s}| < |\Pi_{s p}|$.

  \noindent Such a source $s$ forwards $\alpha_{s}$ to its neighbors.
  Since communication links are reliable, neighboring \processes
  eventually receive such a notification $\forall \langle s, q \rangle
  \in E \iff \eventually r_q(\alpha_{s}^{w_{pq}})$. Most importantly,
  whatever the order of received messages, every \process $q$ in this
  neighborhood such that there exists no closer source $s'$ than the
  received one $s$, delivers it, since $|\Pi_{sq}| < |\Pi_{s'q}|$.

  \noindent They forward it and their respective neighbors eventually
  receive it. By transitivity, the message originating from $s$
  reaches all \processes belonging to $P_s$ at least through their
  shortest paths: $\forall q' \in V, s, s' \in S: |\Pi_{s q'}| <
  |\Pi_{s' q'}| \implies \eventually d_{q'}(\alpha_{s})$.  Since there
  exists only one best sum of weights per \process that can never be
  retracted, \processes eventually reach consistent partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement
forwarding of best to reach consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values piggybacked in messages and
  accumulated over forwardings,
\item \processes only add new partitions to the system,
\item and \processes never join, crash nor leave the system.
\end{inparaenum}
For the sake of simplicity, a \process does not forward a notification
to its sender (Line~\ref{line:accumulator}).  Figure~\ref{fig:add}
illustrates its behavior on a system comprising 4 \processes $a$, $b$,
$c$, and $d$. \Process~$a$ and \Process~$d$ become the sources of
their partition.  They \NAMEB a notification \underline{a}dd message:
$\alpha_a^0$ and $\alpha_d^0$. They initialize their own state with
the lowest possible bound $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, \Process~$b$ receives $\alpha_{d}^{1}$. Since
it improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, \Process~$b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. \Processes $c$ and $d$ will never acknowledge that
Source~$a$ exists. Ultimately, \processes discard last transiting
messages. Despite the obvious lack of traffic optimization, the system
reaches consistent partitioning.

>>>>

\begin{definition}[Autonomous systems]
  An autonomous system is a network comprising nodes and communication
  links that we represent as a \underline{g}raph of
  \underline{v}ertices and \underline{e}dges:
  $G = \langle V, E \rangle$ with $E \in V \times V$. A
  \underline{p}ath $\pi_{xz}$ from \Process~$x$ to \Process~$z$ is a
  sequence of contiguous edges
  $[\langle x, y_1 \rangle, \langle y_1, y_2\rangle, \ldots, \langle
  y_n, z \rangle]$.
\end{definition}

@@

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process~$x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process
  to \process:
  $(b_x(m) \wedge r_y(m)) \implies \exists \pi_{xy}: \forall z \in
  \pi_{xy}, \phi(\mu_z, \sigma_z)$.
\end{definition}

This definition encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}.  It
also highlights that the functioning of epidemic propagation is
well-aligned with the objectives of scoped broadcast. More precisely,
it underlines the transitive relevance of messages, thus a \process
can stop forwarding messages as soon as its predicate becomes
unverified.
%For instance, a \process
%from Paris could scoped broadcast messages to all \processes in
%Paris. This requires \processes to store and maintain their city
%location in their local state. \Processes stop delivering and
%forwarding their received messages when they come from a different
%city.  In another instance, a \process from Paris could scoped
%broadcast messages to all \processes in Paris plus neighboring
%cities. This requires \processes to overload forwarded messages the
%first time they reach another city. The predicate checks if messages
%already reached two distinct cities before delivery. Similarly to
%uniform reliable broadcast, scoped broadcast implementations expose
%different trade-offs on space, time, and communication.
%
We use \NAMEB to efficiently modify the state of each \process
depending on the partitions that exist in the system.

@@

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_{s\in
    S}$ be the \underline{p}artition including at least \Process~$s$,
  each \process belongs to at most one partition $\forall x, y \in V,
  \forall s,s' \in S: (x \in P_{s} \wedge y \in P_{s'}) \implies (x \neq
  y \vee s = s')$, and there exists at least one path $\pi_{xs}$ of
  \processes that belong to this partition $\forall z \in \pi_{xs}: z
  \in P_s$.
\end{definition}

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  We associate each edge $\langle x, y \rangle \in E$ with a pair of
  positive \underline{w}eights $w_{xy}$ and $w_{yx}$. The shortest
  \underline{p}ath $\Pi_{xz}$ has a distance as a sum of weights
  $\sum\{w_{pq} | \langle p, q \rangle \in \Pi_{xz}\} = |\Pi_{xz}|$
  that is smaller than any other path, with $\Pi_{xx}$ being the
  greatest lower bound of $x$. The only consistent partitioning is a
  set of logical partitions $P_{s\in S}$ such that every \process
  belongs to the partition of its closest source:
  $\forall x \in P_{s}: \nexists P_{s'}$ such that
  $|\Pi_{s'x}| < |\Pi_{sx}|$.
\end{definition}

@@

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors:
  $f_x(m) \implies \forall \langle x, y\rangle \in E:
  \TODO{s_{xy}(m)}$.
\end{definition}

\begin{theorem}[\label{theo:efb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually$CP]
    % 
    Assuming reliable communication links
    ($s_{xy}(m) \iff \eventually r_{yx}(m)$), \processes eventually
    reach consistent partitioning if each \process delivers its best
    message among \underline{r}eceived messages $R_x$, and forwards
    its best message among \underline{d}elivered messages $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

@@

\begin{proof}
  % When there are no sources, no \process receives, nor delivers hence
  % forwards any message. \Processes do not belong to any partition.

  \noindent Whenever a \process $s$ becomes a source, it delivers its
  own message $d_s(\alpha_{s})$. Whatever its set of received
  messages, its own partition $P_s$ becomes and remains its closest
  partition $\alpha_{s}^{|\Pi_{ss}|} = \min D_s$ since $\forall x \neq s: |\Pi_{s
    s}| < |\Pi_{s x}|$.
  %
  \noindent Such a source $s$ forwards $\alpha_{s}$ to its neighbors.
  Since communication links are reliable, neighboring \processes
  eventually receive such a notification
  $\forall \langle s, y \rangle \in E \iff \eventually
  r_y(\alpha_{s}^{d})$. Most importantly, whatever the order of
  received messages, every \process $q$ in this neighborhood such that
  there exists no closer source $s'$ than the received one $s$,
  delivers it, since $|\Pi_{sq}| < |\Pi_{s'q}|$.
  %
  \noindent They forward it and their respective neighbors eventually
  receive it. By transitivity, the message originating from $s$
  reaches all \processes belonging to $P_s$ at least through their
  shortest paths: $\forall q' \in V, s, s' \in S: |\Pi_{s q'}| <
  |\Pi_{s' q'}| \implies \eventually d_{q'}(\alpha_{s})$.  Since there
  exists only one best sum of weights per \process that can never be
  retracted, \processes eventually reach consistent partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement
forwarding of best to reach consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values piggybacked in messages and
  accumulated over forwarding,
\item \processes only add new partitions to the system,
\item and \processes never join, crash nor leave the system.
\end{inparaenum}
For the sake of simplicity, a \process does not forward a notification
to its sender (Line~\ref{line:accumulator}).  Figure~\ref{fig:add}
illustrates the behavior of this algorithm on a system comprising 4
\processes $a$, $b$, $c$, and $d$. \Process~$a$ and \Process~$d$
become the sources of their partition.  They \NAMEB a notification
\underline{a}dd message: $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, \Process~$b$ discards $\alpha_{a}^{2}$, for it
does not improve its partition distance. \Processes $c$ and $d$ will
never acknowledge that Source~$a$ exists. Ultimately, \processes
discard last transiting messages. Despite the obvious lack of traffic
optimization, the system reaches consistent partitioning.
